Percussive Sound Generation with Virtual Listeners and Modular Synthesizers

### 2020-01-31 paper.tex
Updates from Overleaf

   \item  $Phase1\,1$: Binary separation of sounds with percussive features from non-percussive sounds.
   \item $Phae2\,2$: Given confidence that the sound being categorized is percussive, categorizing the type of drum or percussion.

@@

\item E+F-DVD: A fully connected model trained on a concatenation of envelope and frequency features for Phase 2. Reaching 80\% accuracy for 6-way drum categorization.
\item CNN-DVD: A deep CNN model that reaches used for Phase 2. 82\% accuracy in a 6-way drum categorization. Trained on Spectrum features.
\item FC-DVD: Fully connected 3 layer neural net with 78\% accuracy for 6-way drum categorization. Trained on Spectrum features.

@@

With our models showing efficacy at categorizing organic drums vs randomly generated synth sounds, and categorizing organic drums accurately, we combine models in order to increase the efficacy of each phase and address the "open set" problem for our task. For Phase 1 we only determine sounds as percussive if both FC-DVN and CNN-LSTM have determined it as such with over 90\% confidence. For the majority of our random generations that is not the case, but if a randomly generated sound is passed this phase, our three categorizers assign their categorizations to this sound. These categorizers have a moderate degree of agree-ability as seen in \ref{survey}, but often there is not an unanimous decision. The Fourth method of categorization, "averaged-cat", is implemented by taking the sum of the softmax outputs of all three categorizers and use this sum to determine the category. \\
These models can be combined and weighted in various ways. The confidence thresholds can be modified in order to implement "ears" with different properties. A glaring issue in the current implementation is the treatment of softmax outputs as a reasonable measure of a models confidence. For various reasons some models tend to have much higher confidence in their scores, skewing attempts at finding a consensus.

% How Can The Synth and Ear Interact?
%
\subsection{How Can The Synth and Ear Interact?}
With our Synth and Ear in place we can quickly generate programs for our Synth, create the corresponding audio sample and get the scores for the audio sample. If the ear returns a vector of categorical probabilities, we can assign categorical ranks to the audio sample. For example, the category with the highest probability will have the rank of 1, second highest would rank 2nd, and the lowest probability will have the rank of n (assuming n categories).\textbf{
But how can we use this to create percussion sounds?}\\\\
We utilize our tools in three different methods of generation, with their specific implementation discussed in section \ref{gens}.
\begin {enumerate} [label=(\roman*)]
\item Random Generation: We randomly create a drum program until a sound belonging to our category is produced. We hypothesize that due to the speed of our experiments it is a viable approach.
\item Multivariate Distribution Sampling: We attempt to learn the joint distribution of our parameters in order to maximize generation of a certain drum category. 
\item Genetic Search: We treat our parameters as genes and attempt to use evolutionary search algorithms to find programs that the ear deems most fit.\\
\end {enumerate} 
\todo[inline,color=blue!30]{As seen in Figure \ref{fig:rank portions}, there is a bias in random generation towards making higher ranking hats and kicks and lower ranking shakers and claps. Is this bias because of Synth structure or a problem with the ear? Will analyse further after updating ear}

\section{Novel Generations}
\label{gens}
\subsection{The Goal}
 Our goal is to rapidly create high quality, novel drum sounds. In this work we analyze random generations in a build-up towards a guided generation via heuristic search.
\subsection{General Methods}
\subsubsection{Random Generation}
To establish a benchmark we seek to assess the likelihood of generating samples of each category if we generate programs for various stack sizes by random selection of parameters. We have observed a bias in generations as indicated in figure \ref{gens}.\\
\subsubsection{Evolutionary Search}
We implemented evolutionary search using DEAP \citep{DEAP_JMLR2012}. A parameter set is assumed to be an individual with each parameter being its genes. To measure an individuals fitness we use the formula $fitness=score-rank$. As generations evolve, a Hall-of-Fame (HOF) list with a fixed size tracks the best individuals to ever live in any given generation, updating itself after each batch of off-springs is evaluated.\\
To ensure diversity of the HOF individuals the fitness score should also consider some measure of uniqueness. A weighted minkowski distance \footnote{https://docs.scipy.org/doc/scipy/reference/generated/scipy.spatial.distance.minkowski.html} between the genes of an offspring and the HOF individuals may be a good measurement.\\
A generation of $N$ individuals are initialized randomly. To create the offspring generation, each generation goes through mutations and in-place gene crossovers. The new generation consists of the top 80\% of the offspring generation and 20\% new individuals to quicken gene diversity. If the new individuals are coming directly from previously measured high scoring parameter sets, we ensure that they cannot be directly added to HOF without going through mutations and crossovers first.\\
\subsubsection{Survey of Generated Drums}
\label{survey}
\begin{figure}[H]
\centering
\includegraphics[width=1.1\linewidth]{images/cat.pdf}
\caption{Frequency of assigned labels for the human categorizers and the average categorizer}
\label{fig:freq-survey}
\end{figure}

To measure the quality of the samples produced by our pipeline and the power of our models, we randomly generated around 50-60 samples in the following categories: "snare","kick","hat","clap" and "other" (combination of rims, 
shakers and unusual percussive sounds"). These samples were determined to be percussive and then categorized by 4 different models (FC, CNNLSTM, E+F and AVG). We ensured a balanced division between samples of stack sizes 1,2 and 4 (each stack is responsible for a third samples under each category). Both writers then categorized these samples without knowledge of other categorizations (human or computation models). It's important to note that:
\begin{itemize}
    \item Humans had an additional category of "bad" for samples that they deemed not percussive. The "bad" category indicates that the sample should have been skipped in phase 1. 
    \item With 6 categorization groups, human's had the same categorization in 47\% of cases
    \item The agreement between the humans and AVG was 44\% and 47\%, not significantly lower than agreement with each other. 
    \item Of 257 samples, both humans agreed with FC, CNNLSTM, AVG and E+F respectively in 78, 76, 76 and 46 of cases.
\end{itemize}
We assess the reliability of agreement between humans and categorization models via the Fleiss' kappa coefficient \cite{fleiss1971measuring}. The value of 0 or less for this coefficient indicates no agreement beyond random chance, and the value of 1 indicating perfect agreement. Our kappa measurements \ref{kappa_table} lie within the 3.5-4.5 range, indicating mild to moderate agreement between humans and machines. We again measure this coefficient after dropping samples that were categorized as "bad" by the authors, as samples that humans deem to be "bad" indicate a failure in phase 1 and arguably should not have been categorized by the models at all. Dropping of samples that both authors deemed "bad" causes 8\% reduction of our data (21 samples) and a small increase in kappa score. Dropping samples either reviewer deemed bad resulted in a 30\% reduction of samples and relatively large increase in kappa scores. \\

@@

\begin{itemize}
    \item Our methodology was successful in generating novel examples of some percussion categories.
    \item Reiterate the importance of ear.
    \item There are potential benefits (e.g: pedagogical, software engineering, efficiency, data search, generality of solution, etc.) to separate some audio generation tasks into multiple parts, rather than throwing everything at some neural net. 
    \item With the right implementation of DSP methods to create a synth, Genetic search (viewed as too slow by previous work) is viable for finding  desired programs and parameters.
    \item Neural nets are function approximators. You don't need them if the actual function is understandable and simple...(but is our task simple enough? how do I prove it?)
    \item When possible, using DSP methods instead of NNs is preferable in many ways(speed,computation,hardware requirements).\\
    \colorbox{green!=30}{Include some speed measurements in appendix?}
    \item ...
\end{itemize}
\section{Future Work}
\begin{itemize}
    \item Potential for other generative methods (e.g reinforcement learning)
    \item Making the synth's sound range more diverse while maintaining speed and efficiency
    \item Can our approach be used for VSTs?
\end{itemize}

\section{Ethical Standards}
We affirm that our work follows the principles of ethical and professional conduct. No drums were harmed in the making of this paper.
\bibliographystyle{abbrv}
    \bibliography{nime-references} 
\end{document}

>>>>

   \item  $Phase1$: Binary separation of sounds with percussive features from non-percussive sounds.
   \item $Phae2$: Given confidence that the sound being categorized is percussive, categorizing the type of drum or percussion.

@@

\item E+F-DVD: A fully connected model trained on a concatenation of envelope and frequency features. Reaching 80\% accuracy for 6-way drum categorization in Phase 2.
\item CNN-DVD: A deep CNN model trained on Spectrum features. Reaching 82\% accuracy in a 6-way drum categorization in Phase 2. 
\item FC-DVD: Fully connected 3 layer neural net with 78\% accuracy for 6-way drum categorization in Phase 2. 

@@

With our models showing efficacy at categorizing organic drums vs randomly generated synth sounds, and categorizing organic drums accurately, we combine models in order to increase the efficacy of each phase and address the "open set" problem for our task. For Phase 1 we only determine sounds as percussive if both FC-DVN and CNN-LSTM have determined it as such with over 90\% confidence. For the majority of our random generations that is not the case, but if a randomly generated sound is passed this phase, our three categorizers assign their categorizations to this sound. These categorizers have a moderate degree of agree-ability as seen in \ref{survey}, but often there is not an unanimous decision. The fourth method of categorization, "averaged-cat", is implemented by taking the sum of the softmax outputs of all three categorizers and use this sum to determine the category. \\
These models can be combined and weighted in various way and the confidence thresholds can be modified in order to implement "ears" with different properties. A glaring issue in the current implementation is the treatment of softmax outputs as a reasonable measure of a models confidence. For various reasons some models tend to have much higher confidence in their scores, skewing attempts at finding a consensus.

\section{Novel Generations}
\label{gens}
\subsubsection{The Pipeline}
 With our Synth and Ear in place we can quickly generate random programs for our Synth, create the corresponding audio sample and get the categorizations for the audio sample. We randomly create drum programs and generate the corresponding signal. This signal is then fed through $Phase 1$ and $Phase 2$ of our ear model to determine if the signal is percussive. If so, a category is assigned to our sound. This sound can then be saved as an audio file. \\
 With our current model, a signal iteration of this pipeline will take approximately 50 milliseconds for up to 8 stacks. This means around 20 iterations can be done per second using a single process. This pipeline scales up efficiently with multiprocessing. We have not measured the iteration latency for this pipeline when a GPU is leveraged. For reference, our CPU measurements are collected on a 2012 Macbook Air running ubuntu 18.04.\\
 Before we attempt a guided generation via genetic or other heuristic search methods, we attempt a deep analysis of the randomly generated samples and the corresponding scores given by the ear. This analysis can expose weaknesses in our methodology which can be improved for future works where various search methods will be attempted.

 \subsubsection{Survey of Randomly Generated Drums}
\label{survey}
\begin{figure}[H]
\centering
\includegraphics[width=1.1\linewidth]{images/cat.pdf}
\caption{Frequency of assigned labels for the human categorizers and the average categorizer}
\label{fig:freq-survey}
\end{figure}
To measure the quality of the samples produced by our pipeline and the power of our models, we randomly generated around 50-60 samples in the following categories: "snare", "kick", "hat", "clap" and "other" (combination of rims, 
shakers and unusual percussive sounds"). These samples were determined to be percussive and then categorized by 4 different models (FC, CNNLSTM, E+F and AVG). We ensured a balanced division between samples of stack size 1, 2 and 4 (each stack is responsible for a third of the samples under each category). Both writers then categorized these samples without knowledge of other categorizations (human or computation models). It's important to note that:
\begin{itemize}
    \item Humans had an additional category of "bad" for samples that they deemed not percussive. The "bad" category indicates that the sample should have been skipped in Phase 1. 
    \item With 6 categorization groups, humans had the same categorization in 47\% of cases.
    \item The agreement between the humans and AVG was 44\% and 47\%, not significantly lower than agreement with each other. 
    \item Of 257 samples humans agreed with FC, CNNLSTM, AVG and E+F respectively in 78, 76, 76 and 46 of cases.
\end{itemize}
We assess the reliability of agreement between humans and categorization models via the Fleiss' kappa coefficient \cite{fleiss1971measuring}. The value of 0 or less for this coefficient indicates no agreement beyond random chance, and the value of 1 indicating perfect agreement. Our kappa measurements \ref{kappa_table} lie within the 3.5-4.5 range, indicating mild to moderate agreement between humans and machines. We again measure this coefficient after dropping samples that were categorized as "bad" by the authors, as samples that humans deem to be "bad" indicate a failure in Phase 1 and arguably should not have been categorized by the models at all. Dropping of samples that both authors deemed "bad" causes an 8\% reduction of our data (21 samples) and a small increase in kappa score. Dropping samples either reviewer deemed bad resulted in a 30\% reduction of samples and relatively large increase in kappa scores. \\

@@

\begin{itemize}
    \item Our methodology was successful in generating novel examples of some percussion categories.
    \item Reiterate the importance of ear.
    \item There are potential benefits (e.g: pedagogical, software engineering, efficiency, data search, generality of solution, etc.) to separate some audio generation tasks into multiple parts, rather than throwing everything at some neural net. 
    \item With the right implementation of DSP methods to create a synth, Genetic search (viewed as too slow by previous work) is viable for finding desired programs and parameters.
    \item Neural nets are function approximators. You don't need them if the actual function is understandable and simple...(but is our task simple enough? how do I prove it?)
    \item When possible, using DSP methods instead of NNs is preferable in many ways (speed, computation, hardware requirements).\\
    \colorbox{green!=30}{Include some speed measurements in appendix?}
    \item ...
\end{itemize}
\section{Future Work}
\begin{itemize}
    \item Potential for other generative methods (e.g reinforcement learning)
    \item Making the synth's sound range more diverse while maintaining speed and efficiency
    \item Can our approach be used for VSTs?
\end{itemize}

\section{Ethical Standards}
We affirm that our work follows the highest principles of ethical and professional conduct. No drums were harmed in the writing of this paper.
\bibliographystyle{abbrv}
    \bibliography{nime-references} 
\end{document}
